The \emph{FrobeniusThresholds} package for \emph{Macaulay2}

### 2019-06-07 Fsing/FThresholdsPaper/FThresholds.tex
minor changes to paper

There are three main sections in this paper, which describe and illustrate the main functionality implemented in the package \emph{FrobeniusThresholds}.
\begin{itemize}
\item[$\circ$] \autoref{sec.Nu} describes the fundamental function {\tt nu}.  As its name suggests, this function is concerned with computing the integers $\nu_I^J(p^e)$ mentioned above.
\item[$\circ$] \autoref{sec.IsFPT} explains the functionality for determining whether a given number is the $F$-pure threshold, or more generally, an \emph{$F$-jumping exponent}.
\item[$\circ$] \autoref{sec.FPT} describes the features of the central function of the package, {\tt fpt}.
\item[$\circ$]  \autoref{sec.FutureDirections} outlines our plans for future improvements to this package.
\end{itemize}
%{\color{red} Move this later}

@@

The last option we describe for {\tt nu} is {\tt IsLocal}.  When the value of {\tt IsLocal} is set to {\tt false} (the default is {\tt true}), then {\tt nu} ignores the option {\tt ContainmentTest} and the input ideal {\tt I}, and instead calculates
\[
\min_{J \in \mathfrak{m}-\Spec R} \nu_I^J(p^e)
\]
% This is done by instead of computing for which $n$ we have $I^{n} \subseteq J^{[p^e]}$ (or equivalently, $(I^n)^{[1/p^e]} \subseteq J$), it 
by determining the values $n$ for which $(I^n)^{[1/p^e]}$ is a unit ideal.

@@

i5 : nu(1, x^3-y^2)

o5 = 5

o5 : QQ
\end{MyVerbatim}
}

@@

The {\tt fpt} function has the option {\tt UseSpecialAlgorithms}, which, when set to {\tt true} (its default value), tells {\tt fpt} to first check whether $f$ is a diagonal polynomial, a binomial, a form in two variables, or a product of factors in simple normal crossing, in that order.

%If $f$ is a monomial, then its $F$-pure threshold is simply the minimum of the reciprocal of the exponent of each variable in $f$.
When $f$ is a diagonal polynomial, a binomial, or a form in two variables, algorithms of Hern\'andez \cite{HernandezFInvariantsOfDiagonalHyp, HernandezFPureThresholdOfBinomial}, or Hern\'andez and Teixeira \cite{HernandezTeixeiraFThresholdFunctions},  are executed to compute the $F$-pure threshold.
Algorithms for computing $F$-pure thresholds of binomials also appear in \cite{ShibutaTakagiLCThresholds}, though they are only valid in certain special characteristics.

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}
i4 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

@@

The above noted algorithm for computing the $F$-pure threshold of a binary form $f$ requires factoring $f$ into linear forms, and {\tt fpt} can sometimes hang when attempting that factorization, when it happens in a Galois field of excessively large order.
This is a situation when the user will want to set the option \texttt{UseSpecialAlgorithms} to \texttt{false}.
However, when a factorization is already known, the user can pass to {\tt fpt} a list containing all the pairwise prime linear factors of $f$, and a list containing their respective multiplicities.

\medspace
{\small
\setstretch{.67}
\begin{MyVerbatim}
i8 : fpt({x, y, x + y, x + 3*y}, {2, 6, 9, 10}) == oo

@@

\section{Possible Future Directions}
\label{sec.FutureDirections}

One natural future direction would be to better incorporate the test ideals $\tau(I^t)$ when computing $F$-thresholds in the case that the ideal $I$ is not necessarily principal.  The theoretical foundation for computing such ideals has already largely been worked out in the paper \cite{SchwedeTuckerTestIdealsOfNonPrincipal}, but such an update to the \emph{FrobeniusThresholds} package would likely the 
\emph{TestIdeals} package to be updated first.

At the moment, the specialized algorithm called by the {\tt fpt} function for computing the $F$-pure threshold of a homogeneous polynomial in two variables limits us to working with the standard grading on the ambient polynomial ring.  Clearly, it is desirable to be able to work with non-standard gradings as well, and results in the paper \cite{HernandezTeixeiraFThresholdFunctions} suggest an algorithm for doing so.  Thus, the package could be improved by finalizing, and implementing, such an algorithm.

Finally, it would be natural to to develop, and implement,  additional algorithms for computing $F$-pure threshold and $F$-jumping numbers in specialized situations apart from monomials, diagonals, binomials, and simple normal crossings.  The first, third, and fourth authors, along with Josep \'Alvarez Montaner and Jack Jeffries,  are currently working on developing such algorithms.  The theoretical foundation of these algorithms lies in polyhedral geometry and integer programming, and as such, should be able to be implemented in \emph{Macaualy2}.

\bibliographystyle{skalpha}
\bibliography{MainBib}

>>>>

The overarching goal of this paper is paper is to describe and illustrate the main functionality implemented in the \emph{FrobeniusThresholds} package.  We break this down into sections, and then conclude by pointing out some potential future directions for the package.

\begin{itemize}
\item[$\circ$] \autoref{sec.Nu} describes the fundamental function {\tt nu}.  As its name suggests, this function is concerned with computing the integers $\nu_I^J(p^e)$ mentioned above.
\item[$\circ$] \autoref{sec.IsFPT} explains the functionality for determining whether a given number is the $F$-pure threshold, or more generally, an \emph{$F$-jumping exponent}.
\item[$\circ$] \autoref{sec.FPT} describes the features of the central function of the package, {\tt fpt}.
\item[$\circ$]  \autoref{sec.FutureDirections} outlines our plans for future improvements to this package.
\end{itemize}
%{\color{red} Move this later}

@@

The last option we describe for {\tt nu} is {\tt IsLocal}.
Notice that $\nu_I^\idealm(p^e)$ can be interpreted as the maximum integer $n>0$ for which $(I^n)^{[1/p^e]} \not\subseteq \idealm$.
When computing its value globally, it is the maximum $n>0$ for which $(I^n)^{[1/p^e]} \neq R$, and when 
the value of {\tt IsLocal} is set to {\tt false} (from its default {\tt true}), then {\tt nu} ignores the option {\tt ContainmentTest} and instead calculates determines this value $n$.

\[
\min_{J \in \mathfrak{m}-\Spec R} \nu_I^J(p^e)
\]
% This is done by instead of computing for which $n$ we have $I^{n} \subseteq J^{[p^e]}$ (or equivalently, $(I^n)^{[1/p^e]} \subseteq J$), it 
by determining the values $n$ for which $(I^n)^{[1/p^e]}$ is a unit ideal.

@@

o4 : QQ
\end{MyVerbatim}
}

@@

The {\tt fpt} function has an option called {\tt UseSpecialAlgorithms}, which, when set to {\tt true} (its default value), tells {\tt fpt} to first check whether $f$ is a diagonal polynomial, a binomial, a form in two variables, or a product of factors in simple normal crossing, in that order.

When $f$ is a diagonal polynomial, a binomial, or a form in two variables, algorithms of Hern\'andez \cite{HernandezFInvariantsOfDiagonalHyp, HernandezFPureThresholdOfBinomial}, or Hern\'andez and Teixeira \cite{HernandezTeixeiraFThresholdFunctions},  are executed to compute the $F$-pure threshold.
Algorithms for computing $F$-pure thresholds of binomials also appear in \cite{ShibutaTakagiLCThresholds}, though they are only valid in certain special characteristics.

i4 : fpt(x^17 + y^20 + z^24) -- a diagonal polynomial

@@

The above noted algorithm for computing the $F$-pure threshold of a binary form $f$ requires factoring $f$ into linear forms, and {\tt fpt} can sometimes hang when attempting that factorization, when the factorization occurs in a Galois field of excessively large order.
This is a situation when the user will want to set the option \texttt{UseSpecialAlgorithms} to \texttt{false}.
However, when a factorization is already known, instead of passing {\tt f} to {\tt fpt}, the user can pass a list containing all the pairwise coprime linear factors of $f$ to {\tt fpt}, and a list containing their respective multiplicities.

i8 : fpt({x, y, x + y, x + 3*y}, {2, 6, 9, 10}) == oo

@@

\newpage
\section{Future Directions}
\label{sec.FutureDirections}

We conclude with some comments on possible improvements to the package.

One natural candidate for improvement would be to extend functionality to better handle nonprincipal ideals.  In particular, the package could be improved if it could use the test ideals $\tau(I^t)$ when $I$ is not necessarily principal.  Fortunately, the theoretical foundation for computing such ideals has already largely been worked out in the paper \cite{SchwedeTuckerTestIdealsOfNonPrincipal}. However, such an improvement would necessarily require an update to the \emph{TestIdeals} package.

\begin{enumerate}
\item Add more cases to {\tt UseSpecialAlgorithms}.
%\item Add the capability to compute non-local {\tt fpt} and non-local {\tt nu} via an option.
\item Write a function that will extract from this sequences $\{ \nu_f^{\idealm}(p^e) \}_{e=1}^{\infty}$ roots of Bernstein-Sato polynomials.  Especially, strata.

\item Extend binary forms functions to handle more general $F$-thresholds (that is, with respect to diagonal ideals).  There is one tricky theoretical things, we don't know when this process would stop.  If we can do this, then we can compute $F$-pure thresholds of binary forms in non-standard gradings.  Cite the paper by Daniel and Pedro \cite{HernandezTeixeiraFThresholdFunctions}.
\end{enumerate}

\newpage
\bibliographystyle{skalpha}
\bibliography{MainBib}
